Domain-specific working memory loads selectively increase negative interpertations of surprised facial expressions

### 2019-09-19 Manuscript/Manuscript.Rmd
reworking introduction

### load necessary libraries ###
suppressPackageStartupMessages(library(readxl)) 
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(openxlsx))
suppressPackageStartupMessages(library(Rmisc))
suppressPackageStartupMessages(library(broom))
```
# Introduction
Facial expressions are rich with affective information, and correctly interpreting these social cues is critical for successfully navigating the social world. Often, facial expressions serve as a clear social signal, but this is not always the case. While a smile likely expresses a positive affective state, other cues are not so clear. For instance, a surprised expression could signal either a positive (e.g., winning the lottery) or negative (e.g., seeing a snake in the woods) affective state in the expresser. When contextual information is limited, individuals differ in their tendency to interpret surprised facial expressions as positive or negative. Importantly, this affective bias extends beyond facial expressions, as individuals often show a similar bias to surprised faces as they do for ambiguous scenes [@neta_neural_2013] or even words (Harp, Petro, Brown & Neta, in prep). This bias towards positive or negative interpretations is known as one's valence bias and myriad factors contribute to one's bias.

@@

@@

Recent work suggests that working memory, while a functional (single?) system, may be distributed and divisible into domain-specific subsystems (e.g., affective working memory; see @mikels_affective_2019). While it is possible to (mostly) isolate perceptual processses (e.g., interpreting emotional facial expressions) under strict laboratory conditions, attention "in the real world" is typically distributed across competing exogenous (e.g., noticing a plane fly above you) and endogenous stimuli (e.g., interoception) (cite). Further, these competing stimuli may be either task-revelevant or task-irrelevant (cite). The literature reports conflicting results for the role of affect on working memory (see @jackson_enhanced_2009 and @mather_emotional_2006), with some suggesting ehnaced recall for emotional stimuli and others reporting impairment in working memory processes. Similarly, working memory also affects emotional processes--converging work suggests that cognitive demands can mitigate negative emotional effects [@dillen_tuning_2008]. Lavie and colleagues -[@lavie_load_2004] suggest a load theory in which cognitive resources are finite, thus limiting the availability for simultaneous processing under large task demands. But, how might concurrent working memory demands affect a more malleable process, like interpreting emotional ambiguity?

Obviously, psychological processes do not occur in a void--that is, it would be quite rare to engage only in a face interpretation task. Instead, people are

Understanding the influences on decision making under ambiguous contexts sheds light on the mechanisms responsible for these individual differences. in the present study we aim to better understand how concurrent task demands (i.e., working memory load) may influence the cognitive resources used to arrive at more top-down driven interpetations of ambiguity, specifically in an emotional context.  
  Recent work suggests that ambiguity resolution in this context requires more cognitive resources/processing compared to clearly valenced faces [@mattek_differential_2016; @neta_dont_2016]. 
The valence bias is trait-like [@neta_corrugator_2009] and generalizes to non-face stimuli [@neta_neural_2013]; however, it is also malleable and may differ depenending on experimental manipulations, including stress inductions or instructions to slow responding [@brown_cortisol_2017; @neta_dont_2016]. Importantly, the valence bias relates to behavior outside of the laboratory; specifically, it is known to relate to depressive symptomology [@petro_initial_2018], at least in children. Chronic negativity biases are common in numerous psychopathologies, including depression and anxiety [@matthews_cognitive_2005].

Distractors and task irrelevant stimuli often have detrimental effects on performance in a variety of tasks [; cite, cite]. Further, domain-specific interference may further exacerbate these effects compared to domain-general stimuli [@gruber_effects_2001]. This effect holds up in the emotional domain; for example, the Stroop task [@stroop_studies_1935] has been modified by some researchers to include emotional stimuli [@whalen_emotional_2006] which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Indeed, neuroimaging work supports the idea that separate systems handle attentional biasing for domain-specific (emotional vs. non-emotional) task relevancy [@egner_dissociable_2008].
Given that a regulatory mechanism likely contributes to positive interpretations of surprised facial expressions, domain-specific interference may cause more negative interpretations of ambiguity compared to a more domain-general interference. Mattek and colleagues [-@mattek_differential_2016] recently showed that different levels of cognitive load (i.e., holding either a single or seven digit number in working memory) does not affect subjective interpretations of surprised facial expressions, but that high cognitive loads do mitigate mouse trajectories. While the authors interpret this as a distinction between trait-like biases and dynamic cognitive-motor processes, there may be more domain-specific processes (e.g., emotional components) that span across these two measures of valence bias. Given the task irrelevance of the numeric distractors in Mattek and colleagues' [-@mattek_differential_2016] work, it follows that the resources required for interpreting ambiguity as positive [@neta_corrugator_2009] may not have been recruited for working memory maintenance, and thus no change in subjective ratings was observed. 
In the present study, we aim to test the effects of low and high working memory loads in both emotional and neutral domains. We expect that trials in which participants are maintaining an emotional working memory load will be more negative than neutral trials. Further, we predict that higher working memory laod trials, specifically in the emotional domain, will result in even more exaggerated negative interpretations. 
<!-- Even simple tasks (e.g., remembering a phone number) reveal the limits of  human cognition. In fact, many researchers believe that there is a magic number of the amount of information able to be held in working memory (7 +- 2 paper). Others have focused on the limits of working memory and successful task performance to inform industrial and organizational policies [@sweller_cognitive_1998]. Despite the large amount of research around working memory and cognitive control, much controversy remains around the mechanisms through which attention to perceptually salient stimuli affect attentional processes and subsequent cognitive control [@lavie_load_2004]. Specifically, there is a divide amongst researchers that believe cognitive loads either inhibit perceptions of non-relevant stimuli -->

>>>>

### load necessary libraries ###
suppressPackageStartupMessages(library(readxl)) 
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(openxlsx))
suppressPackageStartupMessages(library(Rmisc))
suppressPackageStartupMessages(library(broom))
```
# Introduction
Completion of even simple tasks in daily life require regular recruitment of working memory, and this often happens without consideration of how these working memory demands may affect concurrent processes. Many types of information can be held in working memory: strings of numbers, neutral and emotional images, stimulus characteristics like brightness, spatial locations, and more. There is debate in the literature around the neurophysiological substrates of working memory and whether this construct is a unified system or comprised of several domain-specific subsystems (see Baddeley, 1986; Shah & Mikaye, 1996). Recent work suggests though that working memory does indeed contain domain-specific subsystems (e.g., affective working memory; see @mikels_affective_2019)--or at least that interactions between working memory and visual or affective tasks are variable depending on working memory content. Similar domain-specific effects have been observed in a number of other lines of research, including those outside of the working memory domain. For example, the Stroop task [@stroop_studies_1935] has been modified to include emotional stimuli [@whalen_emotional_2006] which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Further, neuroimaging work supports the notion that separate systems handle attentional biasing for domain-specific (emotional vs. non-emotional) task relevancy [@egner_dissociable_2008]. The present work aims to clarify the interaction of working memory demands on a concurrent and ecologically valid emotionoal task (i.e., interpretations of emotional facial expressions), specifically when that content varies in both amount and affective quality.

There is a large body of research concerning the effects of distractor and task irrelevant stimuli on performance in a variety of tasks [cite, cite]. Notably, domain-specific interference may further exacerbate these effects compared to domain-general stimuli [@gruber_effects_2001]. Within an affect-specific context, the literature reports conflicting results for the role of affective stimuli on working memory processing (see @jackson_enhanced_2009 and @mather_emotional_2006), with some suggesting ehnaced recall for emotional stimuli and others reporting impairment in working memory processes. The relationship between these systems is bidirectional though, and often times working memory affects affective processing. One such way is through dampening emotional responses, and converging work suggests that cognitive demands can mitigate negative emotional effects [@dillen_tuning_2008]. One explanation for mitigated emotional responding during concurrent working memory demands is load theory. This theory posits that cognitive resources are finite, thus limiting the availability for simultaneous processing under large task demands [@lavie_load_2004]. One limitation to this previous work is that the affective processes under study involved clearly interpretable stimuli/feelings. A more malleable affective process, such as interpreting ambiguously valenced emotional facial expressions, may offer better understanding of how concurrent working memory demands affect emotional processes.

Facial expressions are rich with affective information, and correctly interpreting these social cues is critical for successfully navigating the social world. Often, facial expressions serve as a clear social signal, but this is not always the case. While a smile likely expresses a positive affective state, other cues are not so clear. For instance, a surprised expression could signal either a positive (e.g., winning the lottery) or negative (e.g., seeing a snake in the woods) affective state in the expresser. When contextual information is limited, individuals differ in their tendency to interpret surprised facial expressions as positive or negative. Importantly, this affective bias extends beyond facial expressions, as individuals often show a similar bias to surprised faces as they do for ambiguous scenes [@neta_neural_2013] or even words (Harp, Petro, Brown & Neta, in prep). This bias towards positive or negative interpretations is known as one's valence bias and myriad factors contribute to one's bias.

@@

Given that a regulatory mechanism likely contributes to positive interpretations of surprised facial expressions, domain-specific interference may cause more negative interpretations of ambiguity compared to a more domain-general interference. Mattek and colleagues [-@mattek_differential_2016] recently showed that different levels of cognitive load (i.e., holding either a single or seven digit number in working memory) does not affect subjective interpretations of surprised facial expressions, but that high cognitive loads do mitigate mouse trajectories. While the authors interpret this as a distinction between trait-like biases and dynamic cognitive-motor processes, there may be more domain-specific processes (e.g., emotional components) that span across these two measures of valence bias. Given the task irrelevance of the numeric distractors in Mattek and colleagues' [-@mattek_differential_2016] work, it follows that the resources required for interpreting ambiguity as positive [@neta_corrugator_2009] may not have been recruited for working memory maintenance, and thus no change in subjective ratings was observed.

In the present study, we aim to test the effects of low and high working memory loads in both emotional and neutral domains. We expect that trials in which participants are maintaining an emotional working memory load will be more negative than neutral trials. Further, we predict that higher working memory laod trials, specifically in the emotional domain, will result in even more exaggerated negative interpretations.

@@

Recent work suggests that ambiguity resolution in this context requires more cognitive resources/processing compared to clearly valenced faces [@mattek_differential_2016; @neta_dont_2016]. 
The valence bias is trait-like [@neta_corrugator_2009] and generalizes to non-face stimuli [@neta_neural_2013]; however, it is also malleable and may differ depenending on experimental manipulations, including stress inductions or instructions to slow responding [@brown_cortisol_2017; @neta_dont_2016]. Importantly, the valence bias relates to behavior outside of the laboratory; specifically, it is known to relate to depressive symptomology [@petro_initial_2018], at least in children. Chronic negativity biases are common in numerous psychopathologies, including depression and anxiety [@matthews_cognitive_2005].

<!-- Distractors and task irrelevant stimuli often have detrimental effects on performance in a variety of tasks [; cite, cite]. Further, domain-specific interference may further exacerbate these effects compared to domain-general stimuli [@gruber_effects_2001]. This effect holds up in the emotional domain; for example, the Stroop task [@stroop_studies_1935] has been modified by some researchers to include emotional stimuli [@whalen_emotional_2006] which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Indeed, neuroimaging work supports the idea that separate systems handle attentional biasing for domain-specific (emotional vs. non-emotional) task relevancy [@egner_dissociable_2008]. -->

### 2019-09-19 Manuscript/Manuscript.tex
reworking introduction

Facial expressions are rich with affective information, and correctly interpreting these social cues is critical for successfully navigating the social world. In fact, facial expressions are seen across cultures ({\textbf{???}}) and some evidence suggests expressions are innate (cite). Often, facial expressions serve as a clear social signal, but this is not always the case. While a smile from a friend likely expresses a positive affective state, other cues are not so clear. For instance, a surprised facial expression could signal either a positive (e.g., winning the lottery) or negative (e.g., a snake in the woods) affective state. In the absence of a larger context, individuals differ in their tendency to interpret surprised facial expressions as either positive or negative. Importantly, this affective bias extends beyond facial expressions, as individuals often show a similar bias to both surprised faces and ambiguous scenes (Neta, Kelley, \& Whalen, 2013). This bias towards positive or negative interpretations is known as one's valence bias.
Interpreting facial expressions requires both bottom-up (e.g., perceptual input) and top-down (e.g., emotion regulation strategies) processes. A growing body of work suggests that the initial interpretation of emotionally ambiguous stimuli is negative and driven by bottom-up processes, and that arriving at a positive interpretation requires additional, top-down regulatory processes. For example, participants reliably rate surprise as negative faster than positive (Neta \& Tong, 2016). In fact, forcing participants to slow their responding during interpretations of ambiguous images shifts individuals' biases towards positivity ({\textbf{???}}). Perceptual input also contributes to valence bias. In one recent study, Neta and colleagues ({\textbf{???}}) showed that faster intial fixation, as well as longer overall fixation, on the mouth is related to more positive interpretations of surprised faces and that forcing gaze patterns to match those of modulated interpretations of surprised expressions. In short,\\
Despite the trait-like nature of this bias (Neta, Norris, \& Whalen, 2009), valence bias may be shifted, at least temporarily, by a number of experimental manipulations. As mentioned above, simple manipulations like slowing response times will shift bias ({\textbf{???}}). Additional work has shown that increases in salivary cortisol after a stressor relate to more negative interpretations of surprised faces from baseline to post-stressor (Brown, Raio, \& Neta, 2017). Similarly, participants with positive biases at baseline will interpret surprise as more negative under threat of shock ({\textbf{???}}), suggesting that taxing cognitive resources, in this case attention, reduces the ability of individuals to interpret surprised faces as positive.

Understanding the influences on decision making under ambiguous contexts sheds light on the mechanisms responsible for these individual differences. in the present study we aim to better understand how concurrent task demands (i.e., working memory load) may influence the cognitive resources used to arrive at more top-down driven interpetations of ambiguity, specifically in an emotional context.\\
Recent work suggests that ambiguity resolution in this context requires more cognitive resources/processing compared to clearly valenced faces (Mattek, Whalen, Berkowitz, \& Freeman, 2016; Neta \& Tong, 2016).
The valence bias is trait-like (Neta et al., 2009) and generalizes to non-face stimuli (Neta et al., 2013); however, it is also malleable and may differ depenending on experimental manipulations, including stress inductions or instructions to slow responding (Brown et al., 2017; Neta \& Tong, 2016). Importantly, the valence bias relates to behavior outside of the laboratory; specifically, it is known to relate to depressive symptomology ({\textbf{???}}), at least in children. Chronic negativity biases are common in numerous psychopathologies, including depression and anxiety ({\textbf{???}}).

Distractors and task irrelevant stimuli often have detrimental effects on performance in a variety of tasks {[}; cite, cite{]}. Further, domain-specific interference may further exacerbate these effects compared to domain-general stimuli (Gruber, 2001). This effect holds up in the emotional domain; for example, the Stroop task (Stroop, 1935) has been modified by some researchers to include emotional stimuli (Whalen, Bush, Shin, \& Rauch, 2006) which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Indeed, neuroimaging work supports the idea that separate systems handle attentional biasing for domain-specific (emotional vs.~non-emotional) task relevancy (Egner, Etkin, Gale, \& Hirsch, 2008).
Given that a regulatory mechanism likely contributes to positive interpretations of surprised facial expressions, domain-specific interference may cause more negative interpretations of ambiguity compared to a more domain-general interference. Mattek and colleagues (2016) recently showed that different levels of cognitive load (i.e., holding either a single or seven digit number in working memory) does not affect subjective interpretations of surprised facial expressions, but that high cognitive loads do mitigate mouse trajectories. While the authors interpret this as a distinction between trait-like biases and dynamic cognitive-motor processes, there may be more domain-specific processes (e.g., emotional components) that span across these two measures of valence bias. Given the task irrelevance of the numeric distractors in Mattek and colleagues' (2016) work, it follows that the resources required for interpreting ambiguity as positive (Neta et al., 2009) may not have been recruited for working memory maintenance, and thus no change in subjective ratings was observed.
In the present study, we aim to test the effects of low and high working memory loads in both emotional and neutral domains. We expect that trials in which participants are maintaining an emotional working memory load will be more negative than neutral trials. Further, we predict that higher working memory laod trials, specifically in the emotional domain, will result in even more exaggerated negative interpretations.

@@

@@

@@

>>>>

Completion of even simple tasks in daily life require regular recruitment of working memory, and this often happens without consideration of how these working memory demands may affect concurrent processes. Many types of information can be held in working memory: strings of numbers, neutral and emotional images, stimulus characteristics like brightness, spatial locations, and more. There is debate in the literature around the neurophysiological substrates of working memory and whether this construct is a unified system or comprised of several domain-specific subsystems (see Baddeley, 1986; Shah \& Mikaye, 1996). Recent work suggests though that working memory does indeed contain domain-specific subsystems (e.g., affective working memory; see Mikels and Reuter-Lorenz (2019))--or at least that interactions between working memory and visual or affective tasks are variable depending on working memory content. Similar domain-specific effects have been observed in a number of other lines of research, including those outside of the working memory domain. For example, the Stroop task (Stroop, 1935) has been modified to include emotional stimuli (Whalen, Bush, Shin, \& Rauch, 2006) which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Further, neuroimaging work supports the notion that separate systems handle attentional biasing for domain-specific (emotional vs.~non-emotional) task relevancy (Egner, Etkin, Gale, \& Hirsch, 2008). The present work aims to clarify the interaction of working memory demands on a concurrent and ecologically valid emotionoal task (i.e., interpretations of emotional facial expressions), specifically when that content varies in both amount and affective quality.

There is a large body of research concerning the effects of distractor and task irrelevant stimuli on performance in a variety of tasks {[}cite, cite{]}. Notably, domain-specific interference may further exacerbate these effects compared to domain-general stimuli (Gruber, 2001). Within an affect-specific context, the literature reports conflicting results for the role of affective stimuli on working memory processing (see ({\textbf{???}}) and ({\textbf{???}})), with some suggesting ehnaced recall for emotional stimuli and others reporting impairment in working memory processes. The relationship between these systems is bidirectional though, and often times working memory affects affective processing. One such way is through dampening emotional responses, and converging work suggests that cognitive demands can mitigate negative emotional effects ({\textbf{???}}). One explanation for mitigated emotional responding during concurrent working memory demands is load theory. This theory posits that cognitive resources are finite, thus limiting the availability for simultaneous processing under large task demands (Lavie, Hirst, Fockert, \& Viding, 2004). One limitation to this previous work is that the affective processes under study involved clearly interpretable stimuli/feelings. A more malleable affective process, such as interpreting ambiguously valenced emotional facial expressions, may offer better understanding of how concurrent working memory demands affect emotional processes.

Facial expressions are rich with affective information, and correctly interpreting these social cues is critical for successfully navigating the social world. Often, facial expressions serve as a clear social signal, but this is not always the case. While a smile likely expresses a positive affective state, other cues are not so clear. For instance, a surprised expression could signal either a positive (e.g., winning the lottery) or negative (e.g., seeing a snake in the woods) affective state in the expresser. When contextual information is limited, individuals differ in their tendency to interpret surprised facial expressions as positive or negative. Importantly, this affective bias extends beyond facial expressions, as individuals often show a similar bias to surprised faces as they do for ambiguous scenes (Neta, Kelley, \& Whalen, 2013) or even words (Harp, Petro, Brown \& Neta, in prep). This bias towards positive or negative interpretations is known as one's valence bias and myriad factors contribute to one's bias.

Both bottom-up (e.g., perceptual input) and top-down (e.g., stereotypes) processes are recruited for interpretation of facial expressions, and a growing body of work suggests that the initial interpretation of emotionally ambiguous stimuli is negative and driven by bottom-up processes. For instance, reaction times are faster for negative interpretations of ambiguous stimuli (just faces??) (Neta \& Tong, 2016). Additionally, presentation of surprised facial expressions as low spatial frequency images, which is processed more readily than high spatial frequency images, biased interpretations towards negativity ({\textbf{???}}). Consequently, under this framework, arriving at a positive interpretation requires additional, top-down regulatory processes and there is evidence to support this as well. For example, forcing participants to slow their responding during interpretations of ambiguous images shifts individuals' biases towards positivity (Neta, Tong, \& Henley, 2018). Neuroimaging evidence supports this initial negativity hypothesis as well, more positive individuals show higher levels of BOLD activation in brain regions recruited during emotion regulation ({\textbf{???}}). In short, slowing response time allows individuals to better regulate the potentially negative information in surprised expressions and to see it in a more positive light.

Given that a regulatory mechanism likely contributes to positive interpretations of surprised facial expressions, domain-specific interference may cause more negative interpretations of ambiguity compared to a more domain-general interference. Mattek and colleagues (2016) recently showed that different levels of cognitive load (i.e., holding either a single or seven digit number in working memory) does not affect subjective interpretations of surprised facial expressions, but that high cognitive loads do mitigate mouse trajectories. While the authors interpret this as a distinction between trait-like biases and dynamic cognitive-motor processes, there may be more domain-specific processes (e.g., emotional components) that span across these two measures of valence bias. Given the task irrelevance of the numeric distractors in Mattek and colleagues' (2016) work, it follows that the resources required for interpreting ambiguity as positive (Neta, Norris, \& Whalen, 2009) may not have been recruited for working memory maintenance, and thus no change in subjective ratings was observed.

In the present study, we aim to test the effects of low and high working memory loads in both emotional and neutral domains. We expect that trials in which participants are maintaining an emotional working memory load will be more negative than neutral trials. Further, we predict that higher working memory laod trials, specifically in the emotional domain, will result in even more exaggerated negative interpretations.

Recent work suggests that ambiguity resolution in this context requires more cognitive resources/processing compared to clearly valenced faces (Mattek et al., 2016; Neta \& Tong, 2016).
The valence bias is trait-like (Neta et al., 2009) and generalizes to non-face stimuli (Neta et al., 2013); however, it is also malleable and may differ depenending on experimental manipulations, including stress inductions or instructions to slow responding (Brown, Raio, \& Neta, 2017; Neta \& Tong, 2016). Importantly, the valence bias relates to behavior outside of the laboratory; specifically, it is known to relate to depressive symptomology ({\textbf{???}}), at least in children. Chronic negativity biases are common in numerous psychopathologies, including depression and anxiety ({\textbf{???}}).

@@

\leavevmode\hypertarget{ref-lavie_load_2004}{}%
Lavie, N., Hirst, A., Fockert, J. W. de, \& Viding, E. (2004). Load theory of selective attention and cognitive control. \emph{Journal of Experimental Psychology: General}, \emph{133}(3), 339--354. doi:\href{https://doi.org/10.1037/0096-3445.133.3.339}{10.1037/0096-3445.133.3.339}

@@

\leavevmode\hypertarget{ref-mikels_affective_2019}{}%
Mikels, J. A., \& Reuter-Lorenz, P. A. (2019). Affective working memory: An integrative psychological construct. \emph{Perspectives on Psychological Science: A Journal of the Association for Psychological Science}, \emph{14}(4), 543--559. doi:\href{https://doi.org/10.1177/1745691619837597}{10.1177/1745691619837597}

@@

\leavevmode\hypertarget{ref-neta_its_2018}{}%
Neta, M., Tong, T. T., \& Henley, D. J. (2018). It's a matter of time (perspectives): Shifting valence responses to emotional ambiguity. \emph{Motivation and Emotion}, \emph{42}, 258--266. doi:\href{https://doi.org/10.1007/s11031-018-9665-7}{10.1007/s11031-018-9665-7}
